---
title: "Lab 1"
authors: "Charlie Jubera"
output: pdf_document
---

```{r warning=FALSE, message=FALSE}
# install.packages('UsingR') 
library(UsingR) 
data(normtemp)
```

## Use the normtemp dataset to answer the following:

### Determine the following statistics for the variable temperature:

```         
    Minimum: 96.30
    
    Maximum: 100.80
    
    Mean: 98.25
    
    Standard Deviation: 0.7331832
```

```{r}
summary(normtemp)
sd(normtemp$temperature)
```

### Does temperature appear to be normally distributed?

Yes, the distribution of temperature looks to be Normally distributed visually.

```{r message=FALSE, warning=FALSE}
# install.packages("ggpubr")
library(ggpubr)
gghistogram(normtemp$temperature) # Use ggpubr gghistogram to speed up graph making
ggqqplot(normtemp$temperature)
```

### Create box plots for temperature. Are there any outliers? Display a reference line at 98.6. Does the median body temperature seem to be 98.6 degrees?

No, the median temperature seems to be lower than 98.6. There does seem to be 3 outliers, two lower and one upper

```{r}
ggboxplot(normtemp$temperature) +
  geom_hline(yintercept = 98.6) +
  coord_flip()
```

## Using the Ameshousing dataset from our in-class examples, run some distributional analysis on Sale_Price, Log(Sale_Price), and Gr_Liv_Area

### Create histograms of these three variables

```{r warning=FALSE}
library(AmesHousing)
ames <- make_ordinal_ames()
gghistogram(ames$Sale_Price, y='density', fill = 'blue', alpha = 0.5) +
  geom_density(alpha = 0.5)
gghistogram(log(ames$Sale_Price), y='density', fill = 'blue', alpha = 0.5) +
  geom_density(alpha = 0.5)
gghistogram(ames$Gr_Liv_Area, y='density', fill = 'blue', alpha = 0.5) +
  geom_density(alpha = 0.5)
```

### Create a QQ Plot for both Sale_Price and Log(Sale_Price). Based on these exploratory procedures, which version of the price information (Sale_Price or Log(Sale_Price)) would you say is closer to being normally distributed?

The log of Sale_Price looks closer to Normal than the unchanged Sale_Price distribution.

```{r}
ggqqplot(ames$Sale_Price)
ggqqplot(log(ames$Sale_Price))
```

## Using the Ameshousing dataset from our in-class examples, determine the following:

### What type of variables are each of these columns (Nominal, Ordinal, or Continuous/Quantitative)? Keep in mind that the way they are represented in the R dataset may not be appropriate, so you should make this determination using your own judgement based on the data you are looking at.

-   Overall_Qual: Ordinal
-   Lot_Shape: Ordinal
-   Heating_QC: Ordinal
-   Lot_Area: Quantitative

```{r}
str(ames)
```
