---
title: "R Notebook"
output: html_notebook
---


```{r}
bike <- read.csv('https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/bike.csv')
```

You want to build a couple of different models and see which one is better. We will learn in later lectures how to do this with test datasets, but for right now we will only do this with training data. First, we need to split the data into training and test. Run the following code to get the training and test split:

```{r}
set.seed(123) 
bike <- bike %>% mutate(id = row_number()) 
train <- bike %>% sample_frac(0.7) 
test <- anti_join(bike, train, by = 'id')
```
What is the observation count in each dataset (train and test)?
train - 12165
test - 5214

You can’t decide which variable is better to predict number of users (cnt), actual temperature (temp) or what the temperature feels like (atemp). You know they are highly correlated (correlation of approximately 0.99), but you feel that each might still provide some valuable information. Maybe the temperature provides (ever so slightly) different information than the feeling temperature. To keep them both in the model, build a ridge regression that has four variables: actual temperature (temp), feeling temperature (atemp), humidity (hum) and wind speed (windspeed). Use CV to find the penalty that minimizes the MSE. What is that lambda value?

7.388165

```{r create dummy train}
train_reg <- train %>% 
  dplyr::select(cnt, temp, atemp, hum, windspeed) %>% 
  mutate_if(is.numeric, ~replace_na(., mean(., na.rm = TRUE))) 

train_x <- model.matrix(cnt ~ ., data = train_reg)[, -1] 
train_y <- train_reg$cnt
```

```{r create dummy test}
test_reg <- test %>% 
  dplyr::select(cnt, temp, atemp, hum, windspeed) %>% 
  mutate_if(is.numeric, ~replace_na(., mean(., na.rm = TRUE))) 

test_x <- model.matrix(cnt ~ ., data = test_reg)[, -1] 
test_y <- test_reg$cnt
```

```{r cv ridge}
library(glmnet) 
bike_ridge <- glmnet(x = train_x, y = train_y, alpha = 0) 


bike_ridge_cv <- cv.glmnet(x = train_x, y = train_y, alpha = 0) 
plot(bike_ridge_cv)
bike_ridge_cv$lambda.min
```
Get predictions for this ridge regression on the test dataset. What are the first 5 predictions from the test dataset from this model? What are the respective actual values of cnt from this dataset?

Predictions:
```{r first 5 predictions}
test$pred <- predict(bike_ridge, s = bike_ridge_cv$lambda.min, newx = test_x)
test$pred %>% head(5)
```
Real:
```{r real}
test$cnt %>% head(5)
```

A MLR with only feeling temperature, humidity, and windspeed had a test dataset MAPE of 478%. What was the MAPE from your ridge regression? From this comparison, which model would you choose?

The MLR has better MAPE so we should use that model.

```{r}
test %>% 
  mutate(ridge_APE = 100*abs((cnt - pred)/cnt)) %>% 
  dplyr::summarise(MAPE_ridge = mean(ridge_APE))
```









